---
title: "Activity Monitoring - Steps"
author: "Michael Conway"
date: "Friday, July 18, 2014"
output: html_document
---
###Loading and preprocessing the data

Imported data via 'read.csv' while formatting the variables: 'steps' = numeric, date = as.Date, and 'interval' = integer. Second step was to create an object omitting NAs. Further formatting was performed at different stages of assignment based on the requirements of the task.

```{r}
amd <- read.csv(file = "amd.csv", header = TRUE, colClasses = c("numeric", as.Date("date",format = "%m/%d/%Y"), "integer" ))
amdn <- na.omit(amd)
```

###Mean steps per day

* Histogram
* Mean and Median
    + Summarize
    + Calculate

```{r}
library(ggplot2)
library(plyr)

ggplot(data = amdn, aes(x = date, y = steps)) + geom_histogram(fill = "red", stat = "identity") + theme(axis.text.x = element_text(angle = 90)) + xlab("Date") + ylab("Steps") + labs(title = "Steps per Day") + theme(plot.title = element_text(size = rel(2), vjust = 2)) + theme(axis.text = element_text(size = rel(.75))) + theme(axis.title = element_text(size = rel(1.5)))

amdnm <- ddply(amdn,~date, summarize, mean=mean(steps),median=median(steps))
amdnm
```

###Average Daily Activity

* Subset by interval and caculate mean
    + Plot mean activity by interval
    + Identify max mean interval 

```{r}

ame <- ddply(amdn,~interval, summarize, mean=mean(steps))

ggplot(data = ame, aes(x = interval, y = mean)) + geom_line(size = 1, colour = "blue") + xlab("Interval(24hrs/5minutes)") + ylab("Average Steps") + labs(title = "Average Steps per Day for October and November") + theme(plot.title = element_text(size = rel(1.5), vjust = 2))+ theme(panel.background = element_rect(fill = "white", colour = "black"))

subset(ame, mean == max(ame$mean))
```

###Imputing Missing Values

* Sum NA's
* NA replacement strategy: use inerval means from 'ame' object to replace NA's in corresponding interval with mean
    + Create column equalling 'steps' then input inverval means into NA's with for loop
  
* Plot histogram
* Calculate mean and median of new data set with replaced NA's

```{r}
sum(is.na(amd$steps))

amd$steps_imp <- amd$steps
for(i in 1:nrow(amd)){
    if(is.na(amd$steps[i])){
        amd$steps_imp[i] <- ame[ame$interval==amd$interval[i], 'mean']
    }
}

ggplot(data = amd, aes(x = date, y = steps_imp)) + geom_histogram(fill = "red", stat = "identity") + theme(axis.text.x = element_text(angle = 90)) + xlab("Date") + ylab("Steps") + labs(title = "Steps per Day") + theme(plot.title = element_text(size = rel(2), vjust = 2)) + theme(panel.background = element_rect(fill = "white", colour = "black")) + theme(axis.title=element_text(size=rel(1.5))) + theme(axis.text = element_text(size = rel(.75)))

amedays <- ddply(amd,~date, summarize, mean=mean(steps_imp),median=median(steps_imp))
amedays
```

*Summary Analysis of impact of imputing inverval means for NA's
    + Means did not change
    + Where days with all NA's replaced with mean intervals, median was 34.11; all other medians per day in both instances of monitoring where zero
    + NA replacement improves graphic appearance without skewing data summary
    
###Weekdays and Weekends

* Create factor with for loop
* Subset by factor and calculate mean
* Create panel plot seperated by factors weekday and weekend
    
```{r}
amd$dates <- as.Date(amd$date)
for(i in 1:nrow(amd)){
    if(weekdays(amd$dates[i]) %in% c("Sunday", "Saturday")) {
        amd$day[i] <- "weekend"
    } else amd$day[i] <- "weekday"
}

amem <- ddply(amd, .(interval, day), summarize, mean=mean(steps_imp))

ggplot(data = amem, aes(x = interval, y = mean)) + geom_line(size = 1, colour = "blue") + xlab("Interval(24hrs/5minutes)") + ylab("Average Steps") + labs(title = "Average Steps per Day for October and November") + theme(plot.title = element_text(size = rel(1.5), vjust = 2))+ theme(panel.background = element_rect(fill = "white", colour = "black")) + facet_grid(day ~ .)


```

